Bayesian Hierarchical Modeling: An Introduction and Reassessment

### 2023-04-06 propriety_section.Rmd
Adding sections with revisions in separate document

header-includes   :
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{setspace}
   - \usepackage{xcolor}
   - \definecolor{mypink}{RGB}{255, 230, 255}
   - \usepackage{todonotes}
   - \newcommand{\Julia}[1]{\todo[inline, color=mypink]{#1}}
   - \newcommand{\Revision}[1]{\textcolor{blue}{#1}}
   - \usepackage{graphicx}
   - \usepackage{float}
   - \usepackage{tabularx}
   - \usepackage{hyperref}

@@

```{r setup, include = FALSE}
library(rmarkdown)
library("papaja")
library(knitr)
library(plyr)
library(dplyr)
library(ggplot2)
library(rstan)
library(brms)
library(LaplacesDemon)
library(gridExtra)
library(ggridges)
library(BayesFactor)
library(crch)
library(foreign)
library(tibble)
library(bridgesampling)
library(cowplot)
library(kableExtra)
library(readr)
library(lattice)
library(Rmisc)
library(devtools)
library(gghalves)
library(bayesplot)
library(ggmcmc)
library(lme4)
library(logspline)
library(truncnorm)
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, fig.pos = "H")
```

\newpage

# Introduction

Imagine a typical scenario in the life of an experimental psychology graduate student: You have just collected a large data set, either online or in the lab. Each participant has spent one valuable hour completing many trials per experimental condition. Now it is time to analyze the data, and your supervisor suggests to run a within-subjects ANOVA. This procedure requires you to first aggregate the data by condition and participant. Doing so reduces your data set from 10,000 rows to just 400. Reducing the data by aggregation makes you a bit queasy: It feels like your efforts in data collection and the participants' time are not valued this way. You know there needs to be a better way. And there is: Hierarchical modeling.

Just as in the scenario above, experiments in psychology are often implemented in a repeated-measures design where participants respond to several items, stimuli, or conditions. From a data analysis perspective, such a design implies that observations are nested within participants. The statistical and the psychological literature agree that the optimal analysis accounting for such a data structure is hierarchical modeling [e.g., @efron1977stein; Lee, -@lee2011cognitive; Rouder & Lu, -@rouder2005introduction]. And yet, it is still not the norm to apply hierarchical modeling to data from psychological experiments. One reason for this is that hierarchical modeling is more difficult than familiar procedures such as ANOVA.

With recent software development, hierarchical modeling has become more and more accessible. In addition, many researchers have argued that hierarchical modeling is easier and the interpretation is more intuitive when it is done in the Bayesian statistical framework [e.g., Lynch, -@lynch2007introduction; Rouder & Lu, 2005; Rouder, Morey, & Pratte, -@rouder2013hierarchical]. 
However, there are challenges to the Bayesian way of hierarchical modeling (Rouder & Lu, 2005). These challenges, such as the choice of priors, the need for model comparisons of highly complex models, and programming skills required to use state-of-the-art software solutions, can make researchers hesitant to use this approach. We believe that existing introductions to Bayesian hierarchical modeling do not adequately address these challenges. For example, they use outdated software (Rouder et al., 2013), do not cover model comparison (Rouder & Lu, 2005), or do not discuss prior specification [@shiffrin2008survey]. This paper aims to make Bayesian hierarchical modeling available to a wider public by providing:

- A comparison of and guidance for the use of two software packages,
- Guidance on prior specification, 
- Practical instructions for Bayes factor model comparison.

Our tutorial is directed at researchers who have a basic understanding of Bayesian inference and parameter estimation and are planning to perform more complex Bayesian analyses.

The application of Bayesian hierarchical analysis will be illustrated using a digit classification task collected by @rouder2005hierarchical. In an experimental task participants repeatedly indicated whether a target digit (i.e., 2, 3, 4, 6, 7, or 8) was greater or smaller than 5. The participants in the task performed multiple trials per condition (i.e., digit). Therefore, observations are nested in participants. Our analyses focus on two effects that can potentially occur in this digit classification task: The digit effect and the side effect. According to the symbolic distance hypothesis [Moyer & Landauer, -@moyer1967time] processing of numbers is analogue. If numbers are closer on the number line, we may confuse them more, and comparing them is more difficult. In line with this hypothesis, the digit effect postulates that response times (RTs) are slower when digits are closer to 5. Additionally, we may hypothesize that RTs are affected by whether the target digit was smaller or greater than 5, yet it is unclear which side of 5 would lead to faster or slower RTs. \textcolor{blue}{The digit classification task is a standard task that usually would be analyzed in a 3 (digit, closest to 5, further away from 5, furthest away from 5) by 2 (side, greater or smaller than 5) within-subjects ANOVA.} In this tutorial, we show how inferences can be improved using a hierarchical Bayesian model instead.

```{r data, include = FALSE, eval = FALSE}
# maybe add object here where I save the link 
indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
colnames(indat)=c('sub','block','trial','stim','resp','rt','error')

hist(indat$rt[indat$sub==0])
table(indat$rt[indat$sub==0])
nrow(indat[indat$sub==0,])

unique(indat$stim)
for (i in unique(indat$stim)){
hist(indat$rt[indat$stim==i])
}

clean=function()
{
  indat=read.table(url('https://raw.githubusercontent.com/PerceptionCognitionLab/data0/master/lexDec-dist5/ld5.all'))
  colnames(indat)=c('sub','block','trial','stim','resp','rt','error')

  bad1=indat$sub%in%c(34,43)
  bad2=indat$rt<250 | indat$rt>2000
  bad3=indat$err==1
  bad4=indat$block==0 & indat$trial<20
  bad5=indat$trial==0

  bad=bad1 | bad2 | bad3 |bad4 |bad5
  dat=indat[!bad,]
  return(dat)
}

# x axis digit 
# y axis: mean RT pp 
indat1 <- clean()

meanrtindat <- aggregate(indat1$rt, list(indat1$stim, indat1$sub), mean)  # mean response time per condition per participant 
meanrt <- aggregate(indat1$rt, list(indat1$stim), mean)
for (i in 1:nrow(meanrtindat)){
  if (meanrtindat$Group.1[i] == 0) {meanrtindat$Group.1[i] <- 2 
  } else if(meanrtindat$Group.1[i] == 1) {meanrtindat$Group.1[i] <- 3 
  } else if(meanrtindat$Group.1[i] == 2) {meanrtindat$Group.1[i] <- 4
  } else if(meanrtindat$Group.1[i] == 3) {meanrtindat$Group.1[i] <- 6 
  } else if(meanrtindat$Group.1[i] == 4) {meanrtindat$Group.1[i] <- 7 
  } else {meanrtindat$Group.1[i] <- 8} 
}

for (i in 1:nrow(meanrt)){
  if (meanrt$Group.1[i] == 0) {meanrt$Group.1[i] <- 2 
  } else if(meanrt$Group.1[i] == 1) {meanrt$Group.1[i] <- 3 
  } else if(meanrt$Group.1[i] == 2) {meanrt$Group.1[i] <- 4
  } else if(meanrt$Group.1[i] == 3) {meanrt$Group.1[i] <- 6 
  } else if(meanrt$Group.1[i] == 4) {meanrt$Group.1[i] <- 7 
  } else {meanrt$Group.1[i] <- 8} 
}
```

```{r figure1code, echo=FALSE, fig.align='left', fig.cap="Average response times in milliseconds in two conditions (i.e., digit 3 and 4). The grey lines represent the individual response time effects, and the blue line the mean aggregated response time effect.", fig.height=4.5, fig.width=6, eval = FALSE}
theme_set(theme_apa(base_size = 9))
meanrtplot <- ggplot(meanrtindat, aes(y = x, x = as.factor(Group.1))) + geom_line(aes(group = Group.2), alpha = .15) + 
  geom_line(data = meanrt, aes(group = 1), color='blue') + geom_point(data = meanrt, aes(group = 1), color='blue') + 
  labs(x = "Digit", y = "Response Time (Ms)") + theme_bw() + theme(axis.line = element_line(colour = "black"), panel.border = element_blank(), panel.background = element_blank())

# meanrtplot

# Difference 3 and 4 
dif34indat <- meanrtindat[which(meanrtindat$Group.1 == 3 | meanrtindat$Group.1 == 4),]
meandif34 <- meanrt[which(meanrt$Group.1 == 3 | meanrt$Group.1 == 4),]

meanrtplot2 <- ggplot(dif34indat, aes(y = x, x = as.factor(Group.1))) + geom_line(aes(group = Group.2), alpha = .15) + geom_point(alpha = .15) + 
  geom_line(data = meandif34, aes(group = 1), color='blue', size = 1) + geom_point(data = meandif34, aes(group = 1), color='blue', shape = 21, size = 2, stroke = 1) + 
  labs(x = "Digit", y = "Response Time (Ms)") + theme_bw() + theme(axis.line = element_line(colour = "black"), panel.border = element_blank(), panel.background = element_blank())

meanrtplot2 
```

```{r figure1, echo=FALSE, fig.align='left', fig.cap="Panel A shows the average response times in seconds for two conditions of the digit classification task (i.e., digits 3 and 4). The grey lines represent the average within individual effects, and the blue line the aggregated effect over all individuals. Panel B shows how individual estimates are shrunken towards the general trend. Sample represents the observed effect per participant in seconds, while Model represents the estimated effects per participant. ", fig.height=3, fig.width=6, out.width ="95%"}
knitr::include_graphics("Images/Fig1_meanrtandhierplot_v2.png")
```

```{r figure1b, echo=FALSE, fig.align='left', fig.cap="RTs per condition (i.e., digit). The grey lines represent the individual RTs, and the blue line the mean aggregated response time.", fig.height=4, fig.width=6, eval = FALSE}

meanrtindat$Group.3 <- ifelse(meanrtindat$Group.1 > 5, 1, 0)
meanside <- with(meanrtindat, tapply(x, Group.3, mean))
meanside2 <- as.data.frame(t(meanside))
meanside3 <- data.frame(y = c(meanside2$`0`, meanside2$`1`), type = c("Smaller", "Larger"))

indside <- with(meanrtindat, tapply(x, list(Group.3, Group.2), mean))
indside2 <- as.data.frame(t(indside))
indside3 <- data.frame(y = c(indside2$`0`,indside2$`1`), type = rep(c("Smaller", "Larger"), each = nrow(indside2)), group = rep(1:52, 2))

# matplot(indside, type = "b", lty = 1, col = adjustcolor(1, .3), pch = 20
#         , ylim = c(300, 1200), xlim = c(0.8, 2.2), xaxt = "n")
# axis(1, 1:2, labels = c("smaller", "larger"))
# lines(meanside, lwd = 2, type = "b", col = "darkred")

level_order <- c("Smaller", "Larger")
theme_set(theme_apa(base_size = 9))
meanrtplot2 <- ggplot(indside3, aes(y = y, x = factor(type, level_order))) + 
  geom_line(aes(group = group), alpha = .15) + geom_point() + 
  geom_line(data = meanside3, aes(group = 1), color='blue', size = 1) + 
  geom_point(data = meanside3, 
             aes(group = 1), color='blue', shape = 21, size = 2, stroke = 1) + 
  labs(x = "", y = "Response Time (Ms)") + 
  theme_bw() + 
  theme(axis.line = element_line(colour = "black"), 
        panel.border = element_blank(), 
        panel.background = element_blank())

# meanrtplot2 
```

## Why Multilevel Modeling?

Before diving into hierarchical modeling, let us consider two alternative approaches: Analyzing aggregated data and analyzing the data separately for each participant. Analyses of aggregated data such as the ANOVA from the scenario above examine a general effect that is assumed to be consistent across subjects. For instance, when investigating whether there is an effect of digit on RT, the mean RT for each participant is computed for each digit-condition, and a repeated-measures ANOVA is conducted on these aggregated scores. Analyses of aggregated data cannot investigate whether a trend is consistent across participants. Therefore, when this option is applied and individual variability is ignored, the general trend can be an inaccurate representation of the true general trend [Rouder, Morey, & Pratte, 2013; Haaf & Rouder, 2017; van Doorn, Aust, Haaf, Stefan, and Wagenmakers, -@van2021bayes]. This is shown in Panel A of Figure \@ref(fig:figure1) where some of the individual trends (i.e., slopes of the grey lines) differ from the general trend (i.e., slope of the blue line). If every participant takes part in each condition, the person variability affects all conditions, thus inducing a correlation across conditions. For example, participants in a digit classification task repeatedly respond to target digits inducing a correlation between the responses to the target digits, such as between target digit 3 and target digit 4. Therefore, the observations are not independent as we assume if we perform analysis on aggregated data. As a result, the type I error rate of a test of the effect increases: there is a higher probability of concluding that there is an effect if there is no effect (Rouder & Lu, 2005).

A second alternative analysis method is to conduct an analysis on each participant's data independently. For example, in the digit classification task, we would estimate one digit effect per person (cf. the grey lines in Figure 1A), but no overall effect (cf. the blue line in Figure 1A). Therefore, when analyzing participant data independently, it is hard to draw any general conclusions.

\Revision{In hierarchical analyses, the general or overall effect is sometimes called fixed effect, and the individual-specific deviations from this effect are in that case referred to as random effects (Rouder \& Lu, 2005). We can also estimate individual effects, not as deviations from an overarching effect, but as effects for each individual, as obtained from a per-participant analysis. Throughout this tutorial, we will use the terms that suit the interpretation of our analyses, namely general effect, individual deviation, and individual effect.}

If estimated for each participant independently, observed individual effects are quite variable because they are perturbed by sample noise. In hierarchical modeling, these individual effects are optimally and automatically corrected towards the general trend [@efron1977stein]. This phenomenon, also called `shrinkage', is illustrated in Panel B of Figure \@ref(fig:figure1). The rightmost observation in the figure is in sample quite large compared to the other observed effects. The estimated effect (green point) is drastically corrected towards the mean. Conversely, the general trend is only influenced slightly by individuals showing a highly divergent trend. By adding the individual variability to the model, we no longer assume independence between observations, but the overall effect in the model leaves the possibility to generalize the effect. Therefore, hierarchical analysis can be considered a middle ground between an independent analysis and an analysis of aggregated data. 
```{r figure2, echo=FALSE, eval = FALSE, fig.cap="Sample represents the observed response time in milliseconds per individual, while model represents the individual response time estimates. Individual estimates are shrunken towards the general trend. ", fig.align = "left", fig.height = 2, fig.width = 3, out.width="70%"}
knitr::include_graphics("Images/hierplotefmor.png")
```

The Bayesian framework offers an intuitive approach for hierarchical modeling (Rouder & Lu, 2005; Lynch, 2007). The flexibility of the Bayesian approach facilitates the implementation of complex hierarchical models (Rouder, Morey, & Pratte, 2013). It also allows for the inclusion of prior knowledge into the analysis, as well as the monitoring of evidence for specific hypotheses as the data accumulate [@wagenmakers2018bayesian]. A complete discussion of the Bayesion approach is beyond the scope of this manuscript. We refer the interested reader to Etz and Vandekerckhove [-@etz2018introduction] and Wagenmakers and colleagues [-@wagenmakers2018bayesian].

# Model Specification 
In the following section, we specify two models that can be used for our application example: a normal and a log-normal model.^[[Online Supplement A](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/A%20-%20Model%20Parameterization%20/Online-Supplement-A---Model-Parameterization-for-Aggregated-and-Individual-Models.pdf) provides the parameterization for models of aggregated and individual data, to illustrate the difference with hierarchical models. ] We are aware that many psychologists do not engage with equations on a daily basis. In this section, however, we try to accessibly summarize the key considerations of model specification. We encourage everyone to read on and not skip the section. Additionally, we provide a non-technical verbal summary.

Throughout the section, we assume that each observation $Y_{ijk}$ is the response time from participant $i$ in trial $k$ of condition $j$. The condition $j$ is the digit presented to the participant, and can take the values 2, 3, 4, 6, 7, and 8. We further assume that the digit side ($x$) is recorded using effect coding where a value of -0.5 indicates that the digit is smaller than 5, and a value of 0.5 indicates that the digit is larger than 5. Additionally, we assume that the digits presented in the trial are dummy-coded in the data using four variables ($u$, $v$, $w$, $z$). If digits 2 and 8 were presented all four variables are zero. The dummy variables represent the digits 3 ($u$), 4 ($v$), 6 ($w$), and 7 ($z$), and each take the value 1 if the respective digit was presented in the trial, and 0 if it was not.

## Normal Model

In the normal model, the observation $Y_{ijk}$ is assumed to come from a normal distribution. The mean of this normal distribution is predicted by the person-specific intercept $\gamma$, the effect of the side (i.e., smaller or greater than 5, $\beta$) and the effect of the distance from 5 (i.e., symbolic distance effect, $\delta$’s). The variance is represented by $\sigma^2$. This is captured in the following equation:

\begin{equation}
Y_{ijk} \sim \text{Normal}(\gamma_{i} + x_j \beta_{i} + u_j \delta_{7i} + v_j \delta_{6 i} + w_j \delta_{4i} + z_j \delta_{3i}, \sigma^2), (\#eq:datalevel)
\end{equation}

where

<!-- \[ -->
<!-- x_j =  -->
<!-- \begin{cases} -->
<!-- \frac{1}{2} & j < 5\\ -->
<!-- - \frac{1}{2} & j > 5, \\ -->
<!-- \end{cases} -->

<!-- \; \; \; -->

<!-- \[ -->
<!-- u_j =  -->
<!-- \begin{cases} -->
<!-- 1 & j = 7\\ -->
<!-- 0 & \text{Otherwise}, -->
<!-- \end{cases} -->

<!-- \; \; \; -->

<!-- v_j =  -->
<!-- \begin{cases} -->
<!-- 1 & j = 6\\ -->
<!-- 0 & \text{Otherwise}, -->
<!-- \end{cases} -->
<!-- \] -->

\begin{align*}
&
\begin{aligned}
x_j = 
\begin{cases}
\frac{1}{2} & j < 5\\
- \frac{1}{2} & j > 5, \\
\end{cases}
\end{aligned}
&
\begin{aligned}
u_j = 
\begin{cases}
1 & j = 7\\
0 & \text{Otherwise},
\end{cases}
\end{aligned}
&
\begin{aligned}
v_j = 
\begin{cases}
1 & j = 6\\
0 & \text{Otherwise},
\end{cases}
\end{aligned}
\end{align*}

<!-- \[ -->
<!-- w_j =  -->
<!-- \begin{cases} -->
<!-- 1 & j = 4\\ -->
<!-- 0 & \text{Otherwise}, -->
<!-- \end{cases} -->

<!-- \; \; \; -->

<!-- z_j =  -->
<!-- \begin{cases} -->
<!-- 1 & j = 3\\ -->
<!-- 0 & \text{Otherwise}. -->
<!-- \end{cases} -->
<!-- \] -->

\begin{align*}
&
\begin{aligned}
w_j = 
\begin{cases}
1 & j = 4\\
0 & \text{Otherwise},
\end{cases}
\end{aligned}
&
\begin{aligned}
z_j = 
\begin{cases}
1 & j = 3\\
0 & \text{Otherwise}.
\end{cases}
\end{aligned}
\end{align*}
\textcolor{blue}{The model equation shows that, for instance, for participant $i = 1$ and digit $j = 7$, with $\gamma_{i = 1}$ equal to $0.5$, $\beta_{i = 1}$ equal to $0.2$, $\delta_{7, i = 1}$ and $\delta_{3, i = 1}$ equal to $0.25$, and $\delta_{6, i = 1}$ and $\delta_{4, i = 1}$ equal to $0.3$, the mean RT is modeled by: $0.5 - 0.1 +  0.25$. If all $\delta$s are positive, then the model equation indicates that individuals respond fastest when the digits 2 and 8 are presented, as in this situation there are no digit effects in the equation. However, for digit $j = 8$, the mean RT is modeled by $0.5 - 0.1$ in this scenario, while for digit $j = 2$, the mean RT is modeled by $0.5 + 0.1$.}

The intercept $\gamma_i$, the side effect $\beta$, and the digit effects $\delta$ are person-specific. This means that for these parameters, the model will produce one estimate for each participant. The multilevel model assumes that these individual-level effects in turn follow a normal distribution:

\begin{equation}
\begin{aligned}
\gamma_{i} &\sim \text{Normal}(\mu_{\gamma}, \sigma^2_{\gamma}), (\#eq:fixedvsrandom)\\
\beta_{i} &\sim \text{Normal}(\mu_{\beta}, \sigma^2_{\beta}), \\
\delta_{7i} &\sim \text{Normal}(\mu_{\delta_{7}}, \sigma^2_{\delta_{7}}), \\
\delta_{6i} &\sim \text{Normal}(\mu_{\delta_{6}}, \sigma^2_{\delta_{6}}), \\
\delta_{4i} &\sim \text{Normal}(\mu_{\delta_{4}}, \sigma^2_{\delta_{4}}),\\
\delta_{3i} &\sim \text{Normal}(\mu_{\delta_{3}}, \sigma^2_{\delta_{3}}).\\
\end{aligned}
\end{equation}
The means ($\mu_{\gamma}$, $\mu_{\beta}$, $\mu_{\delta}$) of these normal distributions represent the general expected effect, for example, $\mu_{\beta}$ represents the expected effect of digit side across individuals. The variances ($\sigma^2_{\gamma}$, $\sigma^2_{\beta}$, $\sigma^2_{\delta}$) represent the individual variation. Therefore, the individual level depends on another level, the overall effect level, containing $\mu_{\gamma}$, $\mu_{\beta}$, $\mu_{\delta}$, $\sigma^2_{\gamma}$, $\sigma^2_{\beta}$, and $\sigma^2_{\delta}$.

## Log-normal Model 
A normal model may not be appropriate in case of more complex theories, hypotheses, or data structures. For example, the data in the example are response times. \Revision{Response times cannot be negative. The normal model, however, can yield negative values. When using the normal model for response times, we are assigning probability to negative values that are not possible. This seems very problematic.} In addition, RT distributions are right-skewed. A better representation of the data at hand is therefore given by a log-normal model [@schramm2019reaction]. This means that Equation \@ref(eq:datalevel) has to be adjusted:

\begin{equation}
Y_{ijk} \sim \text{LogNormal}(\gamma_{i} + x_j \beta_{i} + u_j \delta_{7i} + v_j \delta_{6i} + w_j \delta_{4i} + z_j \delta_{3i}, \sigma^2).  
\end{equation}
Note that the only change in the model formulation is the change of the likelihood function from a normal distribution to a log-normal distribution. \Revision{The likelihood function provides the likelihood of parameter values in the form of probability densities. The log-normal distribution only assigns probability to positive data values, and is, just as the RT distributions, right-skewed. Therefore, the log-normal distribution is a better match for the RT data.} The location parameter of the log-normal distribution is described through the same combination of parameters as the mean of the normal model. However, it has a different interpretation. For example, for a location parameter of $\mu = 0$ and a scale parameter of $\sigma^2 = 1$, the mode of the log-normal distribution is $0.368$, and the mean is $1.65$.

## Summary

In summary, the models are placed directly on the raw RT data. The only difference between the normal and the log-normal models is the likelihood distribution. The normal model just assumes a symmetric normal distribution that can take on any value. The log-normal model assumes a log-normal distribution, which is restricted to positive values and has a right skew.

Both the normal and the log-normal models contain the following parameters: $\gamma_i$ corresponds to the participant's overall response time, $\beta_i$ corresponds to the participant's side effect, and the four $\delta_{\cdot i}$ correspond to the effects between the digits furthest away from 5 (2 or 8) and the other digits. These person-specific effects all come from parent distributions. This results in the hierarchical structure shown in Figure \@ref(fig:hierarchyplot). The figure highlights the relationship between the parameters and the data.

In Bayesian modeling, prior distributions are needed to complete model specification. In the next section, we will take you through the process of setting these priors, followed by sections on Bayesian estimation and hypothesis testing for both the normal and log-normal model. In every section, we will first explain the normal model, followed by an illustration of the log-normal model.

```{r hierarchyplot, echo=FALSE, fig.cap="Hierarchical structure of the normal and log-normal models for the digit classification task.", fig.align = "left", fig.height = 3, fig.width = 4, out.width="70%"}
knitr::include_graphics("Images/Fig2_Hierarchical structure.png")
```

# Prior Settings 
Bayesian analysis requires the specification of prior distributions. Prior distributions are probability distributions on model parameters that specify beliefs about the relative plausibility of parameter values before seeing the data [@wagenmakers2018bayesian]. These prior beliefs can then be updated with the data to obtain the posterior beliefs about the model parameters, following Bayes' rule [@jeffreys1961theory].

For our models, priors are needed on the general effects ($\mu_{\gamma}$, $\mu_{\beta}$, $\mu_{\delta}$) and on the variances ($\sigma_{\gamma}^2$, $\sigma_{\beta}^2$, $\sigma_{\delta}^2$, and $\sigma^2$), as shown in Equation \@ref(eq:fixedvsrandom). For each of these parameters, we specify our beliefs as a probability distribution. For example, if we expect small side effects in the digit classification example, the prior distribution on $\mu_{\beta}$ should have its peak close to zero. Large values for the variance parameters indicate large individual differences, whereas values close to zero indicate that effects are similar across individuals [@haaf2017developing]. If we expect little variability of side effects between people, the prior assigned to $\sigma_{\beta}$ should also have its peak close to zero. The width of prior distributions expresses how certain we are about the parameters before seeing the data. For example, if we want to express that we are certain that the side effect is zero, we can formulate a prior that allocates all probability mass to a parameter value of zero.

## Fear of Commitment  
In many cases substantive researchers fear committing to prior settings because they are unsure about their choices for the type and settings of the distribution. Therefore, they tend to choose wide priors to indicate little prior knowledge about the parameter values [@aczel2018expert]. These include default priors offered by programming packages, such as `rstan` (Stan Development Team, 2019a; illustrated in [Online Supplement B](https://github.com/MyrtheV/Bayesian-Hierarchical-Modelling-An-Introduction-and-Reassessment/blob/main/B%20-%20Default%20Priors%20Rstan/Online-Supplement-B---Default-Priors-Rstan.pdf)) and `brms` [@R-brms_a; @R-brms_b]. Wide priors are priors that allocate (approximately) equal probability density to a great range of values, such as a uniform distribution without bounds [@gelman2006prior]. However, wide priors can be problematic. First, since the distribution is spread out across a wide range of values, the probability density on any specific parameter value is very low. Therefore, any effect is unlikely under the prior as a wide range of effects is deemed plausible. This influences model comparisons: the support for the null hypothesis becomes increasingly large. This is also referred to as the Jeffreys-Lindley paradox [@lindley1957statistical].

Another important issue is that wide priors are often improper, that is, the probability density does not integrate to one [@hobert1996effect]. \Revision{The integral is used as an approximation of the sum for continuous parameters to calculate the probability density.} An example of an improper prior is a uniform prior distribution ranging from minus infinity to infinity. In case of improper priors, an estimation of the general and individual effects cannot be provided. However, even proper priors can result in improper posteriors [@hobert1996effect]. Although estimation should not be possible in this situation, statistical software might still provide results without notifying the user that the posterior distribution does not exist. If users are not aware of this problem, this can result in misleading conclusions. Therefore, it is important to check if *(a)* the priors are proper and *(b)* the priors result in proper posteriors. There is no fail-safe method that guarantees proper posteriors for hierarchical models. The best approach is to choose well-reasoned informative priors for the model, for instance, by using the proposed four-step procedure below.

## Procedure 
We propose four steps to choose suitable priors for Bayesian hierarchical models. We first present the steps and then apply them to choose suitable priors for the parameters $\mu_{\gamma}$, $\mu_{\beta}$, $\mu_{\delta}$, $\sigma_{\gamma}^2$, $\sigma_{\beta}^2$, $\sigma_{\delta}^2$ and $\sigma^2$ in our example.

```{r distributionsteps, echo=FALSE, fig.cap = "Prior specification steps.", fig.align='center', warning=FALSE, message=FALSE, fig.height=3, fig.width=4, cache = TRUE}
# https://jokergoo.github.io/circlize_book/book/graphics.html 
library(circlize)
library(unikn)   # https://cran.r-project.org/web/packages/unikn/vignettes/colors.html

# circos.initialize(letters[1:4], xlim = c(0, 1))
# col = rand_color(4)
# tail = c("point", "normal", "point", "normal")
# circos.track(ylim = c(0, 1), panel.fun = function(x, y) {
#     circos.arrow(x1 = 0, x2 = 1, y = 0.5, width = 0.4, 
#         arrow.head.width = 0.6, arrow.head.length = ux(1, "cm"), 
#         col = col[CELL_META$sector.numeric.index], 
#         tail = tail[CELL_META$sector.numeric.index])
# }, bg.border = NA, track.height = 0.4)

circos.clear()
cell_cycle = data.frame(phase = c("1. Distribution family", "2. Distribution settings", "3. Prior predictions", "4. Adjust"), levels = c("Distribution type", "Distribution settings", "Prior predictions", "Adjust"),
                        hour = c(8, 8, 8, 8))
color = c(NA)
circos.par(start.degree = 90)
circos.initialize(cell_cycle$phase, xlim = cbind(rep(0, 4), cell_cycle$hour))
circos.track(ylim = c(0, 1), panel.fun = function(x, y) {
    circos.arrow(CELL_META$xlim[1], CELL_META$xlim[2], 
        arrow.head.width = CELL_META$yrange*0.8, arrow.head.length = ux(0.5, "cm"),
        col = color[CELL_META$sector.numeric.index])
    circos.text(CELL_META$xcenter, CELL_META$ycenter, CELL_META$sector.index, 
        facing = "bending.inside", niceFacing = TRUE, cex = 0.6)
}, bg.border = NA, track.height = 0.4)
```

## Step 1: Distribution Family  
The first step in specifying a prior distribution is to find a suitable distribution family. For instance, the prior plausibility of parameter values can be expressed through a normal, an inverse gamma, or a Student's *t*-distribution. These distributional families have different shapes and a different support, as illustrated in Figure \@ref(fig:distritypeplot).  For example, the inverse gamma distribution is limited to non-negative parameter values and has a positive skew. The normal and Student's *t*-distribution both have support over the real line, however, the Student's *t*-distribution has wider tails, that is, it assigns a higher plausibility to values that are on the extremes of the distribution.

The support of the prior distribution should cover all theoretically possible parameter values. For example, if a parameter cannot have negative values, such as a standard deviation, the inverse gamma distribution would be a suitable choice. Conversely, if a parameter can take positive and negative values, the normal and Student *t*-distribution would be suitable.

Note that although distributions within a distributional family share the same functional form, they can differ a lot in their shape and location. The exact appearance of a prior distribution is controlled by a set of hyperparameters, for example, the mean and variance for the location and scale of a normal prior. In the next section, we will give some guidance for specifying these hyperparameters.

```{r distritypeplot, echo=FALSE, fig.cap="Common probability density distributions that may be used as priors. Panel A shows normal distributions varying in mean and standard deviation. Panel B shows inverse gamma distributions that differ in the shape and scale parameter. Panel C shows Student's \\textit{t}-distributions that differ in the location and scale parameter. For all three \\textit{t}-distributions, the degrees of freedom are set to 3.", fig.align = "left", fig.height = 8, fig.width = 12, out.width="100%"}
knitr::include_graphics("Images/Fig4_distrtypefigure_v3.png")
```

## Step 2: Set Hyperparameters  
Location, scale, and sometimes shape of the selected distribution can be adjusted by choosing adequate hyperparameters. The goal is to adjust the prior to reflect which parameter values we deem most plausible and how much uncertainty there is. For instance, panel A of Figure \@ref(fig:distritypeplot) illustrates, that increasing the mean in the normal distribution leads to a shift of location: higher parameter values are deemed more plausible according to the pink distribution compared to the green distribution. Panel B shows how uncertainty can be specified, for example, by decreasing the shape of the inverse gamma distribution while keeping the scale constant: the orange distribution assigns probability to a wider range of values than the green distribution.

>>>>

header-includes   :
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{setspace}
   - \usepackage{xcolor}
   - \definecolor{mypink}{RGB}{255, 230, 255}
   - \usepackage{todonotes}
   - \newcommand{\Julia}[1]{\todo[inline, color=mypink]{#1}}
   - \newcommand{\Revision}[1]{\textcolor{blue}{#1}}
   - \usepackage{graphicx}
   - \usepackage{float}
   - \usepackage{tabularx}
   - \usepackage{hyperref}
   - \newenvironment{CSLReferences}{}{\par}

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Section in the Intro about out tutorial:

Our tutorial is directed at researchers who have a basic \Revision{conceptual} understanding of Bayesian inference and parameter estimation and are planning to perform more complex Bayesian analyses \Revision{of psychological phenomena. We hope to achieve a balance of detail and overview---while we want to provide guidance for a full Bayesian workflow (similar to Gelman et al. and Schad et al.) we don't shy away from in-depth explanations for the most curious scientists. To ensure that a broad audience is able to follow the tutorial, we limit and carefully explain equations. However, rudimentary knowledge of calculus is advantageous to fully understand Bayesian analysis.}

### Section about prior propriety

Another important issue is that wide priors are often improper probability distributions [@hobert1996effect]. \Revision{A prior is proper if two conditions hold: 1. All values of the probability distribution are positive definite, and 2. the probability distribution sums to one for discrete data or integrates to one for continuous data. An example of a proper and improper probability distribution are provided in Figure XXX. The figure shows two hypothetical priors for parameter $\theta$. The first condition can be assessed by checking the y-values---both priors have only positive definite values as function output (0.1, 0.2, and 0.4). To assess the second condition, we need to sum up the function values. Panel A shows a proper distribution where the values sum to 1; panel B shows an improper distribution where the values sum to 1.2.}

```{r, echo = F, fig.width=7, fig.height=3.5, fig.align='center', fig.cap="Two hypothetical priors for a discrete parameter $\\theta$. A. A proper probability distribution that sums to one. B. An improper probability distribution that sums to 1.2."}
layout(matrix(1:2, nrow = 1))
par(mgp = c(2, .7, 0))
x <- 1:5
y <- c(12, 6, 6, 3, 3)
y <- y / sum(y)

plot(x, y, pch = 19, col = "gray40", axes = F, ylim = c(0, 1), ylab = "Probability", xlab = expression(theta), main = "A. Proper", xlim = c(.5, 5.5))
axis(1, x)
axis(2, c(0, unique(y), 1), las = 2)
for(i in x) lines(rep(x[i], 2), c(0, y[i]), col = "gray40", lwd = 2)

par(mgp = c(2, .7, 0))
y[3] <- .4

plot(x, y, pch = 19, col = "gray40", axes = F, ylim = c(0, 1), xlim = c(.5, 5.5), ylab = "Probability", xlab = expression(theta), main = "B. Improper")
axis(1, x)
axis(2, c(0, unique(y), 1), las = 2)
for(i in x) lines(rep(x[i], 2), c(0, y[i]), col = "gray40", lwd = 2)
```

A \Revision{more common} example of an improper prior is a uniform prior distribution ranging from minus infinity to infinity. In case of improper priors, \Revision{it is often impossible to obtain correct estimation of general and individual effects in hierarchical models because the resulting posteriors will again be improper probability distributions.} However, even proper priors can result in improper posteriors [@hobert1996effect]. Although estimation should not be possible in this situation, statistical software might still provide results without notifying the user that the posterior distribution does not exist. If users are not aware of the impropriety problem, this can result in misleading conclusions. Therefore, it is important to check if *(a)* the priors are proper and *(b)* the priors result in proper posteriors. \Revision{Unfortunately,} there is no fail-safe method that guarantees proper posteriors for hierarchical models. The best approach is to choose well-reasoned informative priors for the model, for instance, by using the proposed four-step procedure below. \Revision{Priors that are proper and not too diffuse will in most cases lead to proper posteriors.}

### Section on marginal likelihood STILL WORK IN PROGRESS
